*Prochlorococcus marinus* responses to light and oxygen

### 2024-05-24 Output/PLOSRound2/ResponsetoReviewers.txt

L757: The authors should specify that this is a DNA repair ligase, not the DNA ligase that seals Okazaki 
fragments during DNA replication (assuming this is the case).

Thank you for your comment, the revised text now reads "DNA repair ligase".  
Line: XXX

L 758: Photolyase should not have been relevant in the lab experiments because there is no UV light 
emitted (AFAIK) by the bulbs used in the Multicultivator. Also, it is unlikely that the absence of these two 
genes is sufficient to explain why these strains can’t grow at the ocean’s surface.

Thank you. The revised manuscript now reads: "The absence of genes encoding deoxyribodipyrimidine photolyase and DNA ligase (ATP) in P. marinus MIT9313 and P. marinus SS120 may, in part, explain why these two strains cannot tolerate growth under full [O2] and high light, found at the ocean surface."
Line: XXX
XXXXX Mireille to address the wording pertaining to photolyase. We did not imply that photolyase was responsible for the results we found in lab but I will confirm this. XXXXXXXXXX

### 2024-05-27 Code/Manuscript_O2.Rmd
more edits

A Generalized Additive Model (GAM) [@woodGeneralizedAdditiveModels2017]
was applied to the relation of chlorophyll-specific µ, d^-1^ to
photoperiod and PAR level, for each growth [O~2~] level, and for the
blue and red wavebands for growth, for each *P. marinus* strain in this
study. The R package *`mgcv`* [@R-mgcv] was used to model the growth
rate with smoothing terms and indicate the 90, 50 and 10% quantiles for
growth rate across the levels of factors. Only growth rate estimates for
which the amplitude of standard error was smaller than 30% of the fitted
growth rate were included in the GAM. Our priority was the effects of
ecologically relevant blue light on growth trends. We also included GAM
analyses of growth responses to red light, which is not
ecophysiologically relevant, but which might prove mechanistically
informative [@murphyPhotoinactivationPhotosystemII2017].

@@

As with MED4 and SS120, our data again support enhanced growth under
conditions of low cumulative diel PUR and 660 nm (red) spectral
bandwidth, consistent with Murphy *et
al*.[@murphyPhotoinactivationPhotosystemII2017] who found a lower cost
of growth, due to decreased photoinactivation of PSII under red,
compared to blue, wavebands (Fig \@ref(fig:BluevsRedPurFitsPlots)G and H). Interestingly, this protective
effect of red light disappears for MIT9313 growing under 2.5 µM O~2~,
possibly because photoinactivation is strongly suppressed under this low
[O~2~] (Fig \@ref(fig:BluevsRedPurFitsPlots)I). For distinct fits for
different photoperiod fits refer to Fig \@ref(fig:PhotoperiodPurFitsPlots) G-I.

@@

The potential for niche expansion into temperate regions by *P. marinus*
varies depending on the season, which influences achieved underwater
photoperiods and light levels. Temperate summer delivers 11 hours of
blue waveband light underwater, above the photic threshold of 20 µmol
photons m^-2^ s^-1^, while temperate spring/fall delivers 8 hours of
blue waveband light underwater, photoperiod ranges which are permissive
for growth of all three *P. marinus*. In contrast temperate winter
delivers only about 2 h of blue waveband light underwater above the
photic threshold, which precludes growth of MED4 and SS120, even if
winter waters reached permissive temperatures. MIT9313 and SS120 will be
excluded from near-surface growth niches by high PAR, unless OMZ zones
extend to the near surface.

@@

>>>>

A Generalized Additive Model (GAM) [@woodGeneralizedAdditiveModels2017]
was applied to the relation of chlorophyll-specific µ d^-1^ to
photoperiod and PAR level, for each growth [O~2~] level, and for the
blue and red wavebands for growth, for each *P. marinus* strain in this
study. The R package *`mgcv`* [@R-mgcv] was used to model the growth
rate with smoothing terms and indicate the 90, 50 and 10% quantiles for
growth rate across the levels of factors. Only growth rate estimates for
which the amplitude of standard error was smaller than 30% of the fitted
growth rate were included in the GAM. Our priority was the effects of
ecologically relevant blue light on growth trends. We also included GAM
analyses of growth responses to red light, which is not
ecophysiologically relevant, but which might prove mechanistically
informative [@murphyPhotoinactivationPhotosystemII2017].

@@

As with MED4 and SS120, our data again support enhanced growth under
conditions of low cumulative diel PUR and 660 nm (red) spectral
bandwidth, consistent with Murphy *et
al*.[@murphyPhotoinactivationPhotosystemII2017] who found a lower cost
of growth, due to decreased photoinactivation of PSII under red,
compared to blue, wavebands (Fig \@ref(fig:BluevsRedPurFitsPlots)G and H). Interestingly, this protective
effect of red light disappears for MIT9313 growing under 2.5 µM O~2~,
possibly because photoinactivation is strongly suppressed under this low
[O~2~] (Fig \@ref(fig:BluevsRedPurFitsPlots)I). For distinct fits for
different photoperiods refer to Fig \@ref(fig:PhotoperiodPurFitsPlots) G-I.

@@

The potential for niche expansion into temperate regions by *P. marinus*
varies depending on the season, which influences achieved underwater
photoperiods and light levels. Temperate summer delivers 11 hours of
blue waveband light underwater, above the photic threshold of 20 µmol
photons m^-2^ s^-1^, while temperate spring/fall delivers 8 hours of
blue waveband light underwater, photoperiod ranges which are permissive
for growth of all three *P. marinus*. In contrast temperate winter
delivers only about 2 h of blue waveband light underwater above the
photic threshold, which precludes growth of MED4 and SS120, even if
winter waters reached permissive temperatures. MIT9313 and SS120 will be
excluded from near-surface growth niches by high PAR, unless OMZ zones
extend to the near surface (Table \@ref(tab:NichesTable).

@@

XXXX Added this section , if keeping then Mirielle will add the citations to zotero XXXX

The potential expansion of  *P. marinus*  polewards may contribute to the primary food base of an aquatic ecosystem that has already undergone changes in temperate northern waters; commercial fish populations are shifting with new species immigrating  [@zellerCheungZellerPauly2011], endangered whales actively searching for food sources are becoming more prevalent [@meyer-gutbrodMarineSpeciesRange2018] and the white shark seasonal presence is increasing  either due to rising ocean temperatures or following marine prey [@bastienInconspicuousRecoveringNorthward2020]. While the exact impact of the presence of *P. marinus*  on the species distribution of marine wildlife in temperate northern waters is unknown, there is little doubt that *Prochlorococcus* will play a significant role in shaping this distribution.
